Fast radio burst cosmology: dispersion measure angular power spectrum

### 2015-02-05 writeup/dm_angular_ps.tex
Wrote very rough extensions section.

\author{Kris Sigurdson}
 \affiliation{Department of Physics and Astronomy, University of British 
Columbia, Vancouver, BC, Canada V6T 1Z1}

\author{Kiyoshi Wesley Masui}
 \affiliation{Department of Physics and Astronomy, University of British 
Columbia, Vancouver, BC, Canada V6T 1Z1}
 \affiliation{Canadian Institute for Advanced Research, CIFAR Program in
Cosmology and Gravity, Toronto, ON, M5G 1Z8}

@@

@@

@@

@@

@@

>>>>

\author{Kris Sigurdson}
 \affiliation{Department of Physics and Astronomy, University of British 
Columbia, Vancouver, BC, Canada, V6T 1Z1}

\author{Kiyoshi Wesley Masui}
 \affiliation{Department of Physics and Astronomy, University of British 
Columbia, Vancouver, BC, Canada, V6T 1Z1}
 \affiliation{Canadian Institute for Advanced Research, CIFAR Program in
Cosmology and Gravity, Toronto, ON, Canada, M5G 1Z8}

@@

% Dodelson uses \chi not \eta... should we used \chi?

@@

In this section we present a simplified calculation for the sensitivity of the
above estimator which is valid for sample of observed
FRBs distributed roughly uniformly over some fraction of the sky.

The error bar on the redshift averaged dispersion measure angular power spectrum
is
\begin{equation}
    \sigma_{C_l} = \sqrt{\frac{2}{(2 l + 1) f_{\rm sky}}} (C_l + C^N_l),
\end{equation}
where the noise power spectrum is
\begin{equation}
    C^N_l = \frac{4 \pi \sigma_D^2 \sqrt{f_{\rm sky}}}{N_{\rm FRB}}.
\end{equation}
Here $N_{\rm FRB}$ is the number of dispersion measures measured from fast
radio bursts. \red{this depends on the off diagonals of the covariance being
small compared to the diagonals}.
%\red{I'm only 80\% sure about the factors of $f_{\rm sky}$}.

$\sigma_D^2$ is the variance of the dispersion measures.  This would in
practise be measured directly from the observed set, but in terms of
theoretical parameters is
\begin{equation}
\sigma_D^2 = \sigma_{\bar{n}_e}^2 + \sigma^2_{\rm src} + \int \ud z \ud z'
\delta(z - z') p(z) w^{z z'}(1),
\end{equation}
where the first two terms are defined above in equations \ref{e:sig_nebar}
through
\ref{e:sig_src}. The third term is likely
negligible and could also be written as
\begin{equation}
    \int \ud z \ud z' \delta(z - z') p(z) w^{z z'}(1) = \int \ud z \ud z'
    \delta(z - z') p(z) \sum_{l} \frac{2l+1}{4\pi} C_l^{zz}.
\end{equation}

@@

In the above discussions we have focused on the dispersion measure angular
power spectrum due to its conceptual and mathematical simplicity.
However, it is unlikely that this 2D statistic captures all or even most of the
LSS information in a survey.  In this section we discuss possible extensions to
3D statistics that might capture a larger portion of the total available
information.

When discussing the DM angular power spectrum, we treated the dispersion
measure as the \emph{dependant} variable and considered statics. However,
$\Dne$ is in one to one relation to source redshift, and as such the dispersion
measure can be used as a proxy for radial distance.  As such we promote the
dispersion measure to be the \emph{independent} variable and consider the
statistics of the 3D number density field of FRB source events.
If $\Dne$ were the only contribution to the dispersion measure then, assuming
the FRB source events are a biased tracer of matter structure on large scales,
this probe would be broadly similar to galaxy redshift surveys.  Variations in
$\Dsrc$ produce a 'redshift smearing', limiting the effective resolution in the
radial direction. The contributions to the dispersion measure from the electron
over densities, $\Dde$, produce a variety of cosmological signals which we
describe below.

@@

Perturbations in the electron column density due to the intervening large-scale
structure of free electrons results in a mis-estimation of radial distance
from dispersion measure.  We've identified two effects through which this
mis-estimation can cause \emph{unclustered} FRB source events to appear
clustered, which we call the 'dilution effect' and the 'scattering effect'.

In the dilution effect, FRB events embedded in electron over densities have
additional elections (relative to the mean electron density) between them and
thus appear to be more separated along the line of sight and less
clustered.  If both the free electrons and the FRB source events trace the
large-scale structure, this effect is anti correlated with the intrinsic
clustering of the source events, although highly anisotropic.

The scattering effect occurs when perturbations in the foreground free electrons
cause the radial distances to background source events to be mis-estimated. 
When coupled with a non-constant radial distribution
function, $p(\eta)$, this can cause FRB source events to
appear clustered. For instance, if there is an electron over-density in some
direction background FRB events in that direction will be interpreted to be at
a greater radial distance than they are. If $\ud p(\eta) /\ud \eta)$ is
negative, there will be an excess of source events assigned to a given radial
distance behind the electron over density which can be interpreted as apparent
clustering of intrinsically unclustered source events.

Collectively we dub these effects 'dispersion measure space distortions' due
to its broad
similarity to redshift space distortions. In redshift space distortions peculiar
velocities of tracers induce a mis-estimation of radial distance using redshift.
The convergence of these mis-estimations, proportional to the convergence of the
peculiar velocities causes unclustered tracers to appear clustered along the line
of sight. Since the velocity convergence and the over density are perfectly
correlated in linear theory this manifests as an enhancement of line of sight
clustering.

By moving to 3D we fully exploit the redshift information inherent in $\Dne$ as
opposed to treating it largely as a noise term as in our above discussion on
the dispersion measure angular power spectrum. As such we expect such a
treatment would greatly increase the amount of information extracted from an
FRB survey, although we make no effort to quantify this statement.

In addition we expect that in such a treatment the degeneracy between FRB
source clustering and free electron clustering would be significantly less
severe. This is because in 3D, the intrinsic source clustering should be perfectly
isotropic while the electron clustering induces apparent source clustering
only in the radial direction and should thus be highly anisotropic.

@@

In the above dispersion measure space distortions discussion, we discussed how
scattering and dilution can cause unclustered FRB source events to appear to
be clustered in three dimensions.  These effects also have an effect on the
clustered source events. By scattering or diluting the radial distance of
clustered source events, the free electron density field modulates the
statistics of the source clustering. The two point statistics become
non-stationary, which can be measure in the four-point statistics of the source
event density as a non-gaussianity.
